---
title: How to normalize data with Z-scores
   
output: html_document
---

```{r setup, include=F}
# Settings and packages
knitr::opts_chunk$set(echo = TRUE)
```


The most important application of z-scores is in inferential statistics, allowing to calculate p-values for differences between population means(z-test) and sample means (t-test & ANOVA). Displaying data as units of standard deviations can also make visual comparisons across different scales of data possible (e.g. precipitation in $mm$ and temperature in $^\circ{C}$).

### Equation to calculate z-scores for samples

\begin{multline*}
\label{eq:Q1}
z_i = \frac{x_i-\bar{x}}{{S}} 
\end{multline*}

Where:  
$x_i$ is the value of sample $i$\
$\bar{x}$ is the arithmetic mean of the sample\
$S$ is the standard deviation of the sample

### Example in R based on the `iris` data set.

```{r Example}
#Load example data 
data_iris = iris$Sepal.Length[iris$Species=='virginica']

# A vector with 50 numeric values
str(data_iris)

# Manual calculation of z-scores (see equation above)
Z_scores_m = (data_iris - mean(data_iris))/sd(data_iris)
str(Z_scores_m)

# Using the scale() function - default arguments are:  center = TRUE, scale = TRUE
Z_scores_sc = scale(data_iris)
str(Z_scores_m)

#Difference of 0
mean(Z_scores_m -Z_scores_sc)

```

  
